Figures for Paper

### 2021-02-01 paper_figures.Rmd
update git ignore - add some metadata

```{r}
junc_colors = c("#648FFF","#fe6101")
ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))

ipsc_splicing %>% 
  mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ gene_name,
                                    T ~ "")) %>% 
  mutate(junction_name = make.unique(junction_name,sep = ".")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ "")) %>% 
  ggplot(aes(x = deltaPSI, y = log10_test_stat)) +
  geom_point(aes(fill = `Novel Junction`,alpha = graph_alpha), pch = 21, size = 4) + 
  geom_text_repel(aes(label = label_junction), point.padding = 0.3,
                  min.segment.length = unit(0, 'lines')) +
  geom_hline(yintercept = -log10(1 - .9)) + 
  guides(alpha = FALSE, size = FALSE) + 
  scale_fill_manual(values = junc_colors) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab("-Log 10 Test Statistic") + 
  theme(text = element_text(size = 18))
```

# UNC13A Cryptic PSI across TDP-43 KD experiments

@@

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A Cryptic PSI") + 
    xlab("") + 
    guides(color = FALSE)

unca = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A Cryptic PSI") + 
  xlab("") + 
  guides(color = FALSE)

@@

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B Intron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE)

rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    scale_x_discrete(labels=c("shsy5y_normed" = "NB cells \n C/G", 
                              "ipsc_normed_ward" = "iPSC Neurons \n C/G",
                              "shsy5y_normed_no_snp" = "SH-SY5Y \n C/C",
                              "klim_normed" = "Klim iPSC MN \n C/G + C/C",
                              "sh_dzap" = "Appocher \n NB cells \n C/G")) +
    ylab("UNC13A Intron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE)

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
ggarrange(unca,stmn2)
```

# Paient summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi)) %>% 
    mutate(stmn2_psi_groups = as.numeric(cut_interval(log10(stmn_2_cryptic_psi_leaf), n = 2))) %>%
    mutate(stmn2_psi_groups = case_when(stmn2_psi_groups == 1 ~ " Low STMN2",
                                      stmn2_psi_groups == 2 ~ "High STMN2",
                                      TRUE ~ "No STMN2"))

@@

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(disease_group2 = case_when(disease == "Control" ~ "Control",
                                 pathology %in% c("FTD-TDP-A","FTD-TDP-B","FTD-TDP-C") ~ "FTLD-TDP",
                                 pathology %in% c("FTD-TAU","FTD-FUS") ~ "FTLD \n non-TDP",
                                 mutations %in% c("SOD1","FUS") ~ "ALS \n non-TDP",
                                 T ~ "ALS-TDP")) %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel 3'", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel 5'", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel 5'")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```

Looking at disease tissue only, so just taking the cords in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.
```{r}

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals")
```

@@

```{r}
boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected,tissue_clean) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n {tissue_clean} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = disease_group2, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N individuals") + 
  facet_wrap(~tissue_clean) +
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  
```

@@

table_to_test = boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2) %>% 
    unique()

@@

test_pair = test_pair %>% 
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
  mutate(y.position = seq(0.25,by = 0.1,length.out = 9))

>>>>

ipsc_splicing = fread(file.path(here::here(),"data","ipsc_splicing_results.csv"))
ipsc_de = fread(file.path(here::here(),"data","ipsc_differential_expression.csv"))

splicing_dots_tables <- ipsc_splicing %>% 
  mutate(junction_name = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ gene_name,
                                    T ~ "")) %>% 
  mutate(`Novel Junction` = de_novo_junctions == 0) %>% 
  mutate(log10_test_stat = -log10(1 - p_d_psi_0_10_per_lsv_junction)) %>% 
  mutate(log10_test_stat = ifelse(is.infinite(log10_test_stat), 16, log10_test_stat)) %>% 
  mutate(graph_alpha = ifelse(p_d_psi_0_10_per_lsv_junction > 0.9, 1, 0.2)) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2") & 
                                      p_d_psi_0_10_per_lsv_junction > 0.9  & 
                                      deltaPSI > 0 ~ junction_name,
                                    T ~ ""))

fig1a = ggplot() +
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions != 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,,fill = "Annotated Junction"), pch = 21, size = 4) + 
  geom_point(data = splicing_dots_tables %>% filter(de_novo_junctions == 0),
             aes(x = deltaPSI, y =log10_test_stat,
                 alpha = graph_alpha,fill = "Novel Junction"), pch = 21, size = 4) + 
  geom_text_repel(data = splicing_dots_tables[junction_name != ""],
                  aes(x = deltaPSI, y =log10_test_stat,
                      label = label_junction,
                      color = as.character(de_novo_junctions)), point.padding = 0.3,
                    nudge_y = 0.2,
                  min.segment.length = 0,
                  box.padding  = 2,
                  size=6,show.legend = F) +
  geom_hline(yintercept = -log10(1 - .9)) + 
     scale_fill_manual(name = "",
                      breaks = c("Annotated Junction", "Novel Junction"),
                      values = c("Annotated Junction" = "#648FFF", "Novel Junction" = "#fe6101") ) +
       scale_color_manual(name = "",
                      breaks = c("0", "1"),
                      values = c("1" = "#648FFF", "0" = "#fe6101") ) +
  guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab("delta PSI") + 
  ylab(expression(paste("-Lo", g[10], " Test Statistic"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
  xlim(-1,1)

de_table = ipsc_de %>% 
  mutate(contains_cryptic = gene_name %in% splicing_dots_tables[cryptic_junction == T,unique(gene_name)]) %>% 
  mutate(contains_cryptic = as.character(as.numeric(contains_cryptic))) %>% 
  mutate(label_junction = case_when(gene_name %in% c("UNC13A","AGRN",
                                                     "UNC13B","PFKP","SETD5",
                                                     "ATG4B","STMN2","TARDBP") ~ gene_name,
                                    T ~ "")) %>% 
  mutate(graph_alpha = ifelse(padj < 0.1, 1, 0.2)) %>% 
  mutate(y_data = -log10(padj))

fig2a = ggplot(data = de_table) + 
  geom_point(data = de_table %>% filter(contains_cryptic == "0"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "No Cryptic"), pch = 21, size = 4) + 
    geom_point(data = de_table %>% filter(contains_cryptic == "1"),
         aes(x = log2FoldChange, y = -log10(padj),
             alpha = graph_alpha,fill = "Contains Cryptic"), pch = 21, size = 4) + 
    geom_text_repel(data = de_table[label_junction != ""],max.overlaps = 500,
                  aes(x = log2FoldChange, 
                      y = -log10(padj),
                      label = label_junction,
                      color = as.character(contains_cryptic)),
                   nudge_y = 5,
                  min.segment.length = 0,
                  box.padding  = 4,
                  size=6,show.legend = F) +
       scale_fill_manual(name = "",
                      breaks = c("No Cryptic", "Contains Cryptic"),
                      values = c("No Cryptic" = "#648FFF", "Contains Cryptic" = "#fe6101") ) +
         scale_color_manual(name = "",
                      breaks = c("1", "0"),
                      values = c("1" = "#fe6101", "0" = "#648FFF") ) +
  xlim(-7.5,7.5) + 
  ylab(expression(paste("-Lo", g[10], " P-value"))) +
    guides(alpha = FALSE, size = FALSE) + 
  theme(legend.position = 'top') + 
  ggpubr::theme_pubr() + 
  xlab(expression(paste("Lo", g[2], " Fold Change"))) +
  theme(text = element_text(size = 24)) +
  theme(legend.text=element_text(size=22)) +
    geom_hline(yintercept = -log10(0.1))  +
  geom_vline(xintercept=c(0), linetype="dotted")

# UNC13A CE PSI across TDP-43 KD experiments

@@

rel_rna_cryptic_amount = data.table::fread(file.path(here::here(),"data","kd_experiments_relative_rna_and_unc13a_cryptic_junction_counts.csv"))
rel_rna_cryptic_amount[,cryptic_psi_full := ( UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3) / (UNC13A_annotated +  UNC13A_3prime + 
                                          UNC13A_5prime +  UNC13A_5prime_2 + 
                                          UNC13A_5prime_3)]
####barplot UNC13A CE PSI - full five####
unc13a_cryptic_full = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "cryptic_psi_full",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13A CE PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

####barplot UNC13B NMD PSI####
unc13b_psi_nmd = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "unc13b_nmd_exon_psi",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(name = "",
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(name = "",
        values = c("#1C2617","#262114")
    ) +
    ylab("UNC13B \nNMD Exon PSI") + 
    xlab("") + 
    guides(color = FALSE) +
    theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
####UNC13A RNA and UNC13A Cryptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = UNC13A, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
####UNC13A RNA and UNC13A Cryptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = cryptic_psi_full, fill = source),pch = 21,size = 6) + 
    stat_cor(aes(x = TARDBP, y = cryptic_psi_full),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A CE PSI") +
    expand_limits(y = 1) + 
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') +
      theme(text = element_text(size = 18,family = 'sans'), 
          legend.text = element_text(size = 20,family = 'sans'),
          axis.title = element_text(size = 32),
          axis.text = element_text(size = 32))
####UNC13A RNA and UNC13A IR####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = UNC13A, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = UNC13A, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = UNC13A, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom') 
####TARDBP RNA and UNC13A IR####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13a_ir),size = 12) +
    geom_smooth(aes(x = TARDBP, y = cryptic_psi_full),color = "black",method = 'lm') +
    ggpubr::theme_pubr() +
    theme(text = element_text(size = 20)) +
    scale_x_continuous(labels = scales::percent) +
    ylab("UNC13A Normalized IR") +
    theme(legend.title=element_blank()) +
    theme(legend.position = 'bottom')

unca = rel_rna_cryptic_amount %>% 
  ggbarplot(,
            x = "source",
            add = c("mean_se","jitter"),
            y = "cryptic_psi_full",
            fill = 'condition',
            color = 'condition',
            position = position_dodge(0.8)) + 
  ggpubr::theme_pubr() +
  scale_fill_manual(
      values = c("#40B0A6","#E1BE6A")
  ) + 
  scale_color_manual(
      values = c("#1C2617","#262114")
  ) +
  ylab("UNC13A CE PSI") + 
  xlab("") + 
  guides(color = FALSE)

@@

####barplot UNC13B normalized IR####
unc13b_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13b_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13B \nIntron Retention Ratio") + 
    xlab("") + 
    guides(color = FALSE) + 
      theme(legend.position = "none") +
  theme(text = element_text(size = 18)) +
      guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))
####barplot UNC13A normalized IR####
unc13a_ir = rel_rna_cryptic_amount %>% 
    ggbarplot(,
              x = "source",
              add = c("mean_se","jitter"),
              y = "normalized_unc13a_ir",
              fill = 'condition',
              color = 'condition',
              position = position_dodge(0.8)) + 
    ggpubr::theme_pubr() +
    scale_fill_manual(
        values = c("#40B0A6","#E1BE6A")
    ) + 
    scale_color_manual(
        values = c("#1C2617","#262114")
    ) +
    ylab("Normalized UNC13A \nIntron Retention Ratio") + 
    xlab("") + 
    theme(legend.position = "none") +
    guides(color = FALSE) + 
      theme(text = element_text(size = 20,family = 'sans'), 
          legend.text = element_text(size = 36,family = 'sans'),
          axis.title.y = element_text(size = 28),
          axis.text.y = element_text(size = 28))

ggarrange(unc13a_cryptic_full,unc13b_psi_nmd,unc13a_ir,unc13b_ir,common.legend = T)
####scatterplot stmn2 normalized TARDBP####

rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = stmn_2_cryptic_psi, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = stmn_2_cryptic_psi)) +
    theme(legend.position = 'bottom') + 
  facet_wrap(~source)
####scatterplot UNC13BIR normalized TARDBP####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = TARDBP, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = TARDBP, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13AIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13a_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13a_ir)) +
    theme(legend.position = 'bottom')
####scatterplot UNC13BIR and UNC13A Crptic####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = cryptic_psi_full, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = cryptic_psi_full, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)
####scatterplot UNC13BIR and UNC13B NMD####
rel_rna_cryptic_amount %>% 
    filter(condition != "control") %>% 
    ggplot() + 
    geom_point(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir, fill = source),pch = 21,size = 4) + 
    stat_cor(aes(x = unc13b_nmd_exon_psi, y = normalized_unc13b_ir)) +
    theme(legend.position = 'bottom')+ 
  facet_wrap(~source)

####barplot UNC13A RNA level####

# Paient summary statistics
First let's look at only including tissues with detectable stmn2 or UNC13A CE
```{r}
clean_data_table = fread(file.path(here::here(),"data","nygc_junction_information.csv"))
clean_data_table = clean_data_table %>%     
    mutate(call = fct_relevel(call,
                              "C/C", "C/G", "G/G")) %>% 
    mutate(number_g_alleles = as.numeric(call) - 1) %>% 
    mutate(unc13a_cryptic_leaf_psi = ifelse(is.na(unc13a_cryptic_leaf_psi),0,unc13a_cryptic_leaf_psi))

@@

(As a side note, once we do this the number of ALS-non-TDP drops down to 6 (2 FUS) because
the ALS sample tissues are not balanced and not every participant has samples in every tissue)
```{r}
####Inclusion reads by if TDP-potential####
boxplot_table = clean_data_table %>% 
    mutate(across(UNC13A_3prime_leaf:UNC13A_annotated_leaf, ~ .x / library_size,.names = "{.col}_library_norm")) %>% 
    filter(!tissue_clean %in% c("Choroid","Liver")) %>% 
    dplyr::select(sample,participant_id,mutations,disease_group2,pathology,tissue_clean,contains("_library_norm")) %>% 
    melt() %>% 
    filter(grepl("_3prime|_5prime_1",variable)) %>% 
    group_by(sample) %>% 
    mutate(inclusion_reads = sum(value)) %>% 
    ungroup() %>% 
    unique() %>% 
    mutate(junction_name = case_when(variable == "UNC13A_3prime_leaf_library_norm" ~ "  Novel Donor", 
                                 variable == "UNC13A_5prime_1_leaf_library_norm" ~ " Short Novel Acceptor", 
                                 variable == "UNC13A_5prime_2_leaf_library_norm"~ "Long Novel Acceptor")) %>% 
    mutate(disease_tissue = case_when((grepl("FTLD",disease_group2) & grepl("Cortex",tissue_clean))  ~ T,
                                      (grepl("ALS",disease_group2) & grepl("Cord|Motor",tissue_clean))  ~ T,
                                      (grepl("Occipital",tissue_clean)) ~ F,
                                      (grepl("Control",disease_group2) & grepl("Cord|Cortex",tissue_clean)) ~ T,
           TRUE ~ F)) %>% 
    mutate(tissue_clean = gsub("_"," ",tissue_clean))

melt_count = clean_data_table[,.(sample,UNC13A_3prime_leaf,UNC13A_5prime_1_leaf,UNC13A_5prime_2_leaf)] %>% data.table::melt() %>% setnames(.,"value","orig_count")
```
# Detection of UNC13A
Looking at disease tissue only, so just taking the cords in ALS and the 
frontal and temporal cortex of FTLD and then the cord and cortices in Controls.
```{r}

boxplot_table %>% 
  filter(disease_tissue == T) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(participant_id,disease_group2,detected) %>% 
  unique() %>% 
  group_by(disease_group2) %>% 
  mutate(n_sample = n_distinct(participant_id)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate)) + 
  ggpubr::theme_pubr() + 
  scale_y_continuous(lim = c(0,1),labels = scales::percent) + 
  ylab("Percent of Patients \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 24)) + 
  xlab("N individuals")
```

@@

```{r}
boxplot_table %>% 
  filter(!(tissue_clean %in% c("Cerebellum","Hippocampus","Occipital Cortex"))) %>% 
  mutate(detected = inclusion_reads > 0) %>% 
  dplyr::select(sample,disease_group2,detected,tissue_clean) %>% 
  unique() %>% 
  group_by(disease_group2,tissue_clean) %>% 
  mutate(n_sample = n_distinct(sample)) %>% 
  mutate(n_sample_detected = sum(detected)) %>% 
  dplyr::select(tissue_clean,disease_group2,n_sample,n_sample_detected) %>% 
  unique() %>% 
  mutate(detection_rate = n_sample_detected / n_sample) %>% 
  mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
  ungroup() %>% 
  filter(grepl("ALS-TDP|FTLD-T",disease_group2)) %>% 
  mutate(tissue_clean = fct_relevel(tissue_clean, "Cervical Spinal Cord",
                                    "Frontal Cortex",
                                    "Lumbar Spinal Cord",
                                    "Motor Cortex",
                                    "Thoracic Spinal Cord")) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
  mutate(cortex = ifelse(grepl(" Cord",tissue_clean),"cot","cor")) %>% 
  ggplot() +
  geom_col(aes(x = detection_name, y = detection_rate,fill = tissue_clean),position = position_dodge2(width = 0.8, preserve = "single")) + 
  ggpubr::theme_pubr() + 
  ylab("Percent of Tissues \n UNC13A Cryptic Detected") + 
  theme(text = element_text(size = 18)) + 
  xlab("N tissues") + 
  facet_wrap(~tissue_clean,scales = 'free_x',nrow = 3) +
    scale_y_continuous(lim = c(0,1),labels = scales::percent,expand = c(0,0) ) + 
  theme(axis.text.x =  element_text(size = 14)) + 
      theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(text = element_text(size = 18))  +
  scale_fill_manual(values = c("#4C97B5","#C87156","#4C97B5","#C87156","#4C97B5","#C87156")) +
  theme(legend.position = "none") 
```

@@

table_to_test = boxplot_table %>% 
    filter(disease_tissue == T) %>% 
    group_by(disease_group2) %>% 
    mutate(n_sample = n_distinct(sample)) %>% 
    mutate(disease_group2 = gsub("Control"," Control",disease_group2)) %>% 
    mutate(detection_name = glue::glue("{disease_group2} \n ( {n_sample} )")) %>% 
    dplyr::select(detection_name,
                  participant_id,
                  inclusion_reads,
                  disease_group2,
                  tissue_clean,
                  disease_tissue) %>% 
    unique()

@@

test_pair = test_pair %>% 
  mutate(p_value_draw = case_when(p.value < 0.0001~ "***",
                                  p.value < 0.01 ~ "**",
                                   p.value < 0.05 ~ "*",
                          TRUE ~ paste0("Adj. p-value \n",as.character(round(p.value,2))))) %>% 
  mutate(y.position = seq(0.25,by = 0.1,length.out = 7))
